Fractional anisotropy as a tracer of cosmic voids

### 2014-09-04 cosmic_voids.tex
New plots of density profiles of voids for each scheme

%=========================================================================
%		FRONT MATTER
%=========================================================================
\title{Fractional anisotropy as an optimal tracer of cosmic voids}
\author[S. Bustamante and J.E. Forero-Romero]{
\parbox[t]{\textwidth}{\raggedright 
  Sebastian Bustamante \thanks{sebastian.bustamante@udea.edu.co}$^{1}$,
  Jaime E. Forero-Romero$^{2}$ 
}
\vspace*{6pt}\\
$^1$Instituto de F\'{\i}sica - FCEN, Universidad de Antioquia, Calle
67 No. 53-108, Medell\'{\i}n, Colombia\\ 
$^2$Departamento de F\'{i}sica, Universidad de los Andes, Cra. 1
No. 18A-10, Edificio Ip, Bogot\'a, Colombia
}

@@

Since voids where discovered in the first compiled galaxy surveys 
\citep{Chincarini75, Gregory78, Einasto80M, Einasto80N, Kirshner81, 
Kirshner87}, they have been identified, along with the filamentary and 
hierarchically clustered nature of the Cosmic Web, as one of the most 
striking features of the Megaparsec Universe \citep{Bond96}. Nevertheless,
due to the large volume extension of void regions ($\sim 5-10\ \mbox{Mpc} 
h^{-1}$), statistically significant catalogues of voids \citep{Pan10, 
Sutter12b, Nadathur14} only have become available after modern large 
galaxy surveys like the two-degree field Galaxy Redshift Survey \citep{
Colless01, Colless03} and the Sloan Digital Sky Survey \citep{York00,
Abazajian03}, thereby triggering a plethora of more refined observational 
and statistical studies of voids throughout the last decade \citep{Hoyle04,
Croton04, Rojas05, Ceccarelli06, Patiri06a, Tikhonov06, Patiri06b, 
Tikhonov07, BendaBeckmann08, Foster09, Ceccarelli13, Sutter14a}

On the theoretical side, early descriptions of the evolution of the 
large-scale Universe, based on gravitational instabilities in primordial 
stages and leaded by the seminal work of \citet{Zeldovich70}, are highly 
consistent with the Cosmic Web picture, where planar pancake-like regions 
of matter enclose enormous underdense voids and are bordered, in turn, by 
thin filaments and high-density clumpy knots \citep{Bond96}. First 
theoretical models for describing formation, dynamics and properties of 
voids \citep{Hoffman82, Icke84, Bertschinger85, Blumenthal92} were 
dramatically complemented and extended by first numerical studies based 
on simulations \citep{Martel90, Regos91, Weygaert93, Dubinski93}. This 
tendency of using numerical data from N-body simulations, fuelled by last 
generation computing systems and ever more efficient numerical algorithms, 
has become increasingly common in the last years as a powerful analysis 
toolkit of cosmic voids (for an extensive compilation of previous 
numerical works, see \citet{Colberg08}).

At present, studying voids may be considered as a threefold enterprise
\citep{Platen07}: first, they are a key ingredient of the Cosmic Web as 
they dominate almost the entire volume distribution at large-scales and 
additionally, they compensate overdense structures in the total budget of 
matter. This implies that a full understanding of their evolution and 
properties is essential for fathoming the high complexity of the Cosmic 
Web. Second, they provide a valuable resource for inferring and probing 
cosmological parameters as their structure and dynamics are highly 
determined by these values. Third, they constitute and unique and still 
largely pristine environment in which can be tested and probed theories 
of galaxy formation and evolution.

Although a visual recognition of voids on galaxy surveys and simulations
is possible in most cases, a formal systematic identification of them 
is necessary for statistically meaningful studies. However, a basic yet 
essential issue remains regarding the definition itself of what a void is. 
There is not a general consensus about an unambiguous definition of cosmic 
voids and therefore, there can be found many different void finding 
techniques throughout the literature (for a detailed comparison of 
different schemes, see the Void Finder Comparison Project 
\citet{Colberg08}). In spite of the diversity of existing schemes, they 
can be roughly classified into two main types: first, geometric schemes 
based on point spatial or redshift distribution of galaxies in surveys or 
catalogues of dark matter halos in simulations \citep{Kauffmann91, 
Muller00, Gottlober03, Hoyle04, Brunino07, Foster09, Micheletti14, 
Sutter14}, and second, schemes based on the smooth and continuous matter 
density field either from simulations or from reconstruction procedures 
on surveys \citep{Plionis02, Colberg05, Shandarin06, Platen07, Neyrinck08,
Neyrinck13, Ricciardelli2013}.

@@

With the above parameters the mass of each particle in the simulation is 
$m_{\rm p}=1.4\times 10^{8}$\hMsun. Throughout this paper we shall use a
catalogue of halos identified through the Bound Density Maximum algorithm
\citep{Klypin99}. These halos have been selected to have an overdensity of
200 times the critical density in addition to be mutually excluded of the
virial radius of other halos. Mass, comoving position and peculiar 
velocity of each halo at $z=0$ were obtained through the public available 
Multidark database \footnote{\url{http://www.multidark.org/MultiDark/}} 
\citep{Riebe11}. For estimating the density and velocity fields we use a 
\textit{cloud-in-cell} (\texttt{CIC}) algorithm onto a grid of $256^3$ 
cells, corresponding to a resolution of $0.98$\hMpc per cell side. Then 
through finite-differences and FFT methods the tidal and shear tensors 
are computed. Finally the eigenvalues and eigenvectors of each tensor are
obtained for each cell of the grid. Neglecting substructures presented 
below Megaparsec scales and taking into account our focus in voids, which
are a prominent characteristic of the Megaparsec Universe, we apply a 
Gaussian softening of one cell to all fields.

@@

Although the density field is a key quantity for understanding the 
large-scale Universe, the filamentary structure of the Cosmic Web is 
poorly traced by it (see e.g. figure \ref{fig:FA_field}). In order to 
compare our new proposal about the the fractional anisotropy as a tracer 
field of voids, we define a very basic approach for environmental finding 
based on three thresholds of the density field.

@@

%.........................................................................
%FIGURE 3: volume functions
\begin{flushleft}
\begin{figure*}
\centering

@@

>>>>

%=========================================================================
%		FRONT MATTER
%=========================================================================
\title{Fractional anisotropy as a tracer of cosmic voids}
\author[S. Bustamante and J.E. Forero-Romero]{
\parbox[t]{\textwidth}{\raggedright 
  Sebastian Bustamante \thanks{sebastian.bustamante@udea.edu.co}$^{1}$,
  Jaime E. Forero-Romero$^{2}$ 
}
\vspace*{6pt}\\
$^1$Instituto de F\'{\i}sica - FCEN, Universidad de Antioquia, Calle
67 No. 53-108, Medell\'{\i}n, Colombia\\ 
$^2$Departamento de F\'{i}sica, Universidad de los Andes, Cra. 1
No. 18A-10, Edificio Ip, Bogot\'a, Colombia
}

@@

Since voids where discovered in the first compiled galaxy surveys 
\citep{Chincarini75, Gregory78, Einasto80M, Einasto80N, Kirshner81, 
Kirshner87}, they have been identified, along with the filamentary and 
hierarchically clustered nature of the Cosmic Web, as one of the most 
striking features of the Megaparsec Universe \citep{Bond96}. Nevertheless,
due to the large volume extension of void regions ($\sim 5-10\ \mbox{Mpc} 
h^{-1}$), statistically meaningful catalogues of voids \citep{Pan10, 
Sutter12b, Nadathur14} only have become available after modern large 
galaxy surveys like the two-degree field Galaxy Redshift Survey \citep{
Colless01, Colless03} and the Sloan Digital Sky Survey \citep{York00,
Abazajian03}, thereby triggering a plethora of more refined observational 
and statistical studies of voids throughout the last decade \citep{Hoyle04,
Croton04, Rojas05, Ceccarelli06, Patiri06a, Tikhonov06, Patiri06b, 
Tikhonov07, BendaBeckmann08, Foster09, Ceccarelli13, Sutter14a}

On the theoretical side, early descriptions of the evolution of the 
large-scale Universe, based on gravitational instabilities in primordial 
stages and leaded by the seminal work of \citet{Zeldovich70}, are 
consistent with the Cosmic Web picture, where planar pancake-like regions 
of matter enclose enormous underdense voids and are bordered, in turn, by 
thin filaments and high-density clumpy knots \citep{Bond96}. First 
theoretical models for describing formation, dynamics and properties of 
voids \citep{Hoffman82, Icke84, Bertschinger85, Blumenthal92} were 
dramatically complemented and extended by first numerical studies based 
on simulations \citep{Martel90, Regos91, Weygaert93, Dubinski93}. This 
tendency of using numerical data from N-body simulations, fuelled by last 
generation computing systems and ever more efficient numerical algorithms, 
has become increasingly common in the last years as a powerful analysis 
toolkit of cosmic voids (for an extensive compilation of previous 
numerical works, see \citet{Colberg08}).

At present, studying voids may be considered as a threefold enterprise
\citep{Platen07}: first, they are a key ingredient of the Cosmic Web as 
they dominate almost the entire volume distribution at large-scales and 
additionally, they compensate overdense structures in the total budget of 
matter. This implies that a full understanding of their evolution and 
properties is essential for fathoming the high complexity of the Cosmic 
Web. Second, they provide a valuable resource for inferring and probing 
cosmological parameters as their structure and dynamics are highly 
determined by these values. Third, they constitute and unique and still 
largely pristine environment in which can be tested galaxy evolution.

Although a visual recognition of voids in galaxy surveys and simulations
is possible in most cases, a formal systematic identification is necessary 
for statistically meaningful studies. However, a basic yet essential issue 
remains regarding the definition itself of what a void is. There is not a 
general consensus about an unambiguous definition of cosmic voids and 
therefore, there can be found many different void finding techniques 
throughout the literature (for a detailed comparison of different schemes, 
see the Void Finder Comparison Project \citet{Colberg08}). In spite of the 
diversity of existing schemes, they can be roughly classified into two 
main types: first, geometric schemes based on point spatial or redshift 
distribution of galaxies in surveys or catalogues of dark matter halos in 
simulations \citep{Kauffmann91, Muller00, Gottlober03, Hoyle04, Brunino07, 
Foster09, Micheletti14, Sutter14}, and second, schemes based on the smooth 
and continuous matter density field either from simulations or from 
reconstruction procedures on surveys \citep{Plionis02, Colberg05, 
Shandarin06, Platen07, Neyrinck08, Neyrinck13, Ricciardelli2013}.

@@

With the above parameters the mass of each particle in the simulation is 
$m_{\rm p}=1.4\times 10^{8}$\hMsun. Throughout this paper we shall use a
catalogue of halos identified through the Bound Density Maximum algorithm
\citep{Klypin99}. These halos have been selected to have an overdensity of
200 times the critical density in addition to be mutually excluded of the
virial radius of other halos. Mass, comoving position and peculiar 
velocity of each halo at $z=0$ were obtained through the public available 
Multidark database \footnote{\url{http://www.multidark.org/MultiDark/}} 
\citep{Riebe11}. For estimating the density and velocity fields we use a 
\textit{cloud-in-cell} (\texttt{CIC}) algorithm onto a grid of $256^3$ 
cells, corresponding to a resolution of $0.98$\hMpc per cell side. Then, 
through finite-differences and FFT methods the tidal and shear tensors 
are computed. Finally the eigenvalues and eigenvectors of each tensor are
obtained for each cell of the grid. Neglecting substructures presented 
below Megaparsec scales and taking into account our focus in voids, which
are a prominent characteristic of the Megaparsec Universe, we apply a 
Gaussian softening of one cell to all fields.

@@

Although the density field is a key quantity for understanding the 
large-scale Universe, the filamentary structure of the Cosmic Web is 
poorly traced by it (see e.g. figure \ref{fig:FA_field}). In order to 
compare our new proposal based on the the fractional anisotropy as a 
tracer field of voids, we define a very basic approach for environmental 
finding based on three thresholds of the density field.

@@

%.........................................................................
%FIGURE 4: volume functions
\begin{flushleft}
\begin{figure*}
\centering

@@

%.........................................................................
%FIGURE 5: Density profile of voids for each defined scheme
\begin{flushleft}
\begin{figure*}
\centering
  \includegraphics[trim = 3mm 2mm 4mm 0mm, clip, keepaspectratio=true,
  width=0.24\textheight]{./figures/voids_density_TwebFAG.pdf}
  \includegraphics[trim = 3mm 2mm 4mm 0mm, clip, keepaspectratio=true,
  width=0.24\textheight]{./figures/voids_density_VwebFAG.pdf}
  \includegraphics[trim = 3mm 2mm 4mm 0mm, clip, keepaspectratio=true,
  width=0.24\textheight]{./figures/voids_density_TwebDLG.pdf}

  \captionof{figure}{\small Density profiles of voids for each finding 
  scheme.}
